Creating optimal conditions for reproducible data analysis in \R with `fertile'\protect\thanks{The authors gratefully acknowledge contributions from Hadley Wickham, Jenny Bryan, Greg Wilson, Edgar Ruiz, and other members of the \pkg{tidyverse} team.}

### 2020-07-19 STAT-paper-template/APA/fertile.Rnw
added experiment description

To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020, comparing one group of students that received the package functions in their \R environment to one that did not. Unfortunately, however, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing.

We hope to continue this experimentation and gather data on \pkg{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.

>>>>

To test \pkg{fertile}'s effectiveness, we began an initial randomized control trial of the package on an introductory undergraduate data science course at Smith College in Spring 2020. Unfortunately, however, the trial had to be postponed after the COVID-19 pandemic significantly altered the experimental conditions at the midpoint of testing. \\

The experiment was structured as follows:

\begin{itemize}

\item Students are given a form at the start of the semester asking whether they consent to participate in a study on data science education. In order to successfully consent, they must provide their system username, collected through the command \cmd{Sys.getenv("LOGNAME")}. Since these individuals each username belongs to are often easily identifiable, each email address is then transformed into a hexadecimal string via a hashing algorithm in to maintain privacy.

\item These hexadecimal strings are then randomly assigned into equally sized groups, one experimental group that receives the features of \pkg{fertile} and one group that receives a control.

\item The students are then asked to download a package titled the same as the course number of the class: \pkg{sds192}, which was custom created for the purpose of this trial. It leverages an \cmd{.onAttach} function to scan the R environment and collect the username of the user who is loading the package and run it through the same hashing algorithm as used previously. It then identifies whether that user belongs to the experimental or the control group. Depending on the group they are in, they receive a different version of the package on loading:

\item The experimental group receives the basic \pkg{sds192} package, which consists of some data sets and RMarkdown templates necessary for completing homework assignments and projects in the class, but they also have \pkg{fertile} installed and loaded silently in the background. The package's proactive features are enabled without their knowledge, and they can receive warning messages when they use absolute or non-portable paths or attempt to change their working directory. The control group receives only the basic \pkg{sds192} package, including its data sets and RMarkdown templates. They do not have \pkg{fertile} installed and loaded in the background and do not receive proactive warning messages. All students from both groups then use their version of the package throughout the semester on a variety of projects.

\item In order to test \pkg{fertile}'s effectiveness, both groups are given a short quiz on different components of reproducibility that are intended to be taught by \pkg{fertile} at both the beginning and end of the semester. Their scores are then compared to see whether one group learned more than the other group or whether their scores were essentially equivalent.

\end{itemize}

Although the experiment was unsuccessful in its first attempt, we hope to run the same trial again and gather data on \pkg{fertile}'s effectiveness once the higher education environment returns to a relatively stable state.
